A Critical Appraisal of Zero-Multiple Method in Classical Water Simulation

### 2014-11-03 writedown/zm/paper.water/main.tex
backup the work

The equilibrium density of the system was determined by an NPT SPME
simulation.  In the NVT simulation the SPME electrostatic, computed
pressure ($-0.4\pm 1.5$~Bar) is reproducing the target pressure of
1.0~Bar.  The accuracy of ZM method, can be investigate by checking if
the NVT pressure is consistent with the target pressure.  In
Fig.~\ref{fig:pres-comp}, the pressure calculated by ZM method of
order $l=1,2,3$ and 4 are plotted against the cut-off radius.
All the  simulations are done with system $\systemla$.
It is concluded from the Figure that all ZM methods converges to the
target pressure at large enough cut-off radius.
ZM $l=1$ converges from an under-estimate of the pressure, while ZM $l=2,3$ and 4 converges
from an over-estimate.
The higher order of ZM presents an larger error in the pressure calculation than
the lower order ZM method. The ZM method is also compared with
the RF method using the experimental dielectric constant $\varepsilon_{\textrm{rf}} = 80$, 
from which it is concluded that the accuracy of ZM in pressure calculation is much higher
than the RF method.
It should be noted that even with the largest system of 13824 water molecules,
the fluctuation (measured by standard deviation) of the pressure is roughly 100~Bar, which means that it is
very difficult achieve very accurate estimate for this quantity. In practice a pressure error of 20~Bar is
still acceptable.

@@

Ohter thermodynamic quantities including the chemical potential,
constant volume/pressure molar heat capacity, isothermal
compressibility and heat expansion coefficient are compared in
Tab.~\ref{tab:thermo}.
The ZM results are in good consistency with the SPME results, and out-perform
the RF method.

\recheck{Discussion of the energy correction in the chemical potential calculation. TO be added later}

@@

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt/fig-rdf.eps}  
  \caption{The comparison of the center-of-mass radial distribution functions
    calculated by SPME, ZM and RF methods. The insert is a zoom-in of
    range 0.9 -- 1.5~nm. In the insert, the position of the cut-off
    (1.2~nm) is indicated by a vertical black line.
    For ZM method, the splitting parameter $\alpha$ was set to 0.
  }
  \label{fig:rdf}
\end{figure}

@@

\recheck{Shall we comment here on the fact that group cut-off produces smoother RDF?}

@@

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt.small/fig-gkr-small.eps}
  \includegraphics[]{fig/result.nvt.small/fig-gkr-small-conv.eps}
  \includegraphics[]{fig/result.nvt.small/fig-eps-t-3000.eps}
  \includegraphics[]{fig/result.nvt.small/fig-eps-t-zm3-3000.eps}
  \includegraphics[]{fig/result.nvt.small/fig-eps-t-10000.eps}
  \includegraphics[]{fig/result.nvt.small/fig-eps-t-zm3-10000.eps}
  \caption{The Kirkwood G-factor calculated for system $\systemmb$.
    % The result of different methods are compared in the left plot.
    For all ZM methods the splitting parameter $\alpha = 0.00\,\textrm{nm}^{-1}$.
    In the left plot the cut-off radius for all orders is set to $1.2$~nm, while
    in the right plot, only the order $l=3$ is considered and different cut-off radii are compared.
    The RF method uses a cut-off radius of 1.2~nm, and dielectric constant of 80.
    The statistical uncertainty of the SPME method is presented at 95~\% confidence level with the red bar.
  }
  \label{fig:tmp1}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt/fig-gkr.eps}
  \includegraphics[]{fig/result.nvt/fig-gkr-conv.eps}
  \includegraphics[]{fig/result.nvt/fig-eps-t-3000.eps}
  \includegraphics[]{fig/result.nvt/fig-eps-t-zm3-3000.eps}
  \caption{The Kirkwood G-factor calculated for system $\systemlb$.
    % The result of different methods are compared in the left plot.
    For all ZM methods the splitting parameter $\alpha = 0.00\,\textrm{nm}^{-1}$.
    In the left plot the cut-off radius for all orders is set to $1.2$~nm, while
    in the right plot, only the order $l=3$ is considered and different cut-off radii are compared.
    The RF method uses a cut-off radius of 1.2~nm, and dielectric constant of 80.
    The statistical uncertainty of the SPME method is presented at 95~\% confidence level with the red bar.
  }
  \label{fig:tmp2}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt/fig-gkr-10000.eps}
  \includegraphics[]{fig/result.nvt/fig-gkr-conv-10000.eps}
  \includegraphics[]{fig/result.nvt/fig-eps-t-10000.eps}
  \includegraphics[]{fig/result.nvt/fig-eps-t-zm3-10000.eps}
  \caption{The Kirkwood G-factor calculated for system $\systemlb$. Equilibriation time is 10~ns, and the production simulation time is 10~ns.
    % The result of different methods are compared in the left plot.
    For all ZM methods the splitting parameter $\alpha = 0.00\,\textrm{nm}^{-1}$.
    In the left plot the cut-off radius for all orders is set to $1.2$~nm, while
    in the right plot, only the order $l=3$ is considered and different cut-off radii are compared.
    The RF method uses a cut-off radius of 1.2~nm, and dielectric constant of 80.
    The statistical uncertainty of the SPME method is presented at 95~\% confidence level with the red bar.
  }
  \label{fig:tmp3}
\end{figure}

\subsection{Damping effect}

In this section we investigate the damping effect. The pressure is plotted as a function of  the cut-off radius is plotted for different choices of
the splitting parameter $\alpha$ in Fig.~\ref{fig:damp-pres}. The system for the simulation was $L_0$.
\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{fig/nvt.pressure.1/pressure-l1.eps}
  \includegraphics[width=0.49\textwidth]{fig/nvt.pressure.1/pressure-l2.eps}\\
  \includegraphics[width=0.49\textwidth]{fig/nvt.pressure.1/pressure-l3.eps}
  \includegraphics[width=0.49\textwidth]{fig/nvt.pressure.1/pressure-l4.eps}
  \caption{The pressure convergence with respect to the cut-off radius.
    For ZM method, different splitting parameter $\alpha$ (in unit of $\textrm{nm}^{-1}$) are shown for different orders:  $l=1$, 2, 4, and 4.
  }
  \label{fig:damp-pres}
\end{figure}

The convergence of the integrated auto-correlation function $I_\eta(T) $ is shown in Fig.~\ref{fig:damp-vis-l1}.
\begin{figure}
  \centering
  \includegraphics[]{fig/result.tiny/fig-vis-l1-damp.eps}
  \caption{The convergence of the integrated auto-correlation function $I_\eta(T) $. The error bars indicating 95\% confidence level are plotted with the SPME method. The different splitting parameters for ZM $l=1$ method are plotted}
  \label{fig:damp-vis-l1}
\end{figure}

The RDF is plotted for different splitting parameters $\alpha$ for ZM $l=2$ and 3 in Fig.~\ref{fig:damp-rdf-l23}. Higher $\alpha$ reproduces the RDF better.
The system was $L_1$.
\begin{figure}[]
  \centering
  \includegraphics[width=0.49\textwidth]{fig/result.nvt/fig-rdf-l2-damp.eps}
  \includegraphics[width=0.49\textwidth]{fig/result.nvt/fig-rdf-l3-damp.eps}
  \caption{Radial distribution function for different splitting parameters $\alpha$ for ZM $l=2$ (left) and $l=3$ (right).}
  \label{fig:damp-rdf-l23}
\end{figure}

The Kirkwood G-factor is plotted for different splitting parameters $\alpha$ for ZM $l=2$ and 3 in Fig.~\ref{fig:damp-gkr-l23}. 
The system was $L_1$.
\begin{figure}[]
  \centering
  \includegraphics[width=0.49\textwidth]{fig/result.nvt/fig-gkr-l2-damp.eps}
  \includegraphics[width=0.49\textwidth]{fig/result.nvt/fig-gkr-l3-damp.eps}
  \caption{Kirkwood G-factor for different splitting parameters $\alpha$ for ZM $l=2$ (left) and 3 (right).}
  \label{fig:damp-gkr-l23}
\end{figure}

It is well know that the G-factor is system size
dependent~\cite{vanderSpoel2006origin}. Here we show the G-factor for
both a smaller system $\systemmb$ (Fig.~\ref{fig:tmp1}) and a larger system $\systemlb$
(Fig.~\ref{fig:tmp2}).  In the figures, the statistical uncertainty of
the SPME method is denoted by the red error bar, while those of the
other methods are essentially the same, so they are not presented for
clarity.  In both the system, the RF method is qualitatively
wrong. The results of the ZM method are consistent with SPME
method. The lower order produces better consistency.  An artificial
oscillation around the cut-off radius presents for $l=1$, but it is
not obvious for $l\geq 2$.  For high order ZM (here $l=3$ is investigated),
using larger cut-off radius can substantially improve the accuracy.

The
resulting slop is used to calculate the diffusion constant. The results are listed in Table~\ref{tab:thermo}.
The ZM method $l\geq 2$ is consistent with the SPME
result, while the RF method is off.

@@

\subsection{Radial distribution function}
The radial distribution function (RDF) $g(r)$ is a scalor function of distance $r$, which indicates the
probability of finding two atoms of distance $r$ apart.  It is a very
impotant equilibrium structure property, by which the X-ray scattering
intensity and a mount of thermodynamic properties can be computed.
In this work, we investigate the center-of-mass RDF in the water
system.

@@

\subsection{Pressure}
The pressure of the system calculated by the virial formulus:
\begin{align}
  P = \frac1{3V}
  \Bigg\langle
  \sum_{i=1}^{3N} \Big( m_i\vect v_i^2 + \vect r_i\cdot \vect F_i \Big)
  \Bigg\rangle,
\end{align}
where $V$ is the volume of the system.

@@

The chemical potential is defined by
\begin{align}
  \mu = \Big(\frac{\partial A}{\partial N}\Big)_{V,T}
\end{align}
In practice the excess chemical potential $\mu^\exc$, which is the
chemical potential abstracted by the kinetic contribution, is of
special interest. We calculate the excess chemical potential by the
thermodynamic integration (TI). In this approach, the interaction
between an inserted testing molecule and the system is
denotedd by $U_t(\lambda)$, where $\lambda$
is a coupling parameter in range $[0,1]$.
When $\lambda = 1$ the testing molecule is fully coupled to the system
with the TIP3P interaction. When $\lambda=0$, the testing molecule
is decoupled with the system.
Therefore the
Hamiltonian of the system is function of the coupling parameter. The
excess chemical is calculated by the free energy difference between
the fully coupled and decoupled states:
\begin{align}
  \mu^\exc = \int_0^1 \Big\langle \frac{\partial U_t(\lambda)}{\partial \lambda} \Big\rangle_\lambda d\lambda
\end{align}
This integral is usually calculated by numerical integration, which
equally paritions the range of $\lambda$, simulates the system at each
discretized $\lambda$, and computes chemical potential by trapezoidal formulus.
Here for the sake of accuracy, 
we adopt a two step coupling approach: firstly couple the van der Waals
interaction of the testing particle with the rest of the system (stage vdw),
and then couple the electrostatic interaction (stage ele).
During stage vdw, 21 $\lambda$ values are equally
distributed in $[0,1]$.  During stage ele, 6 $\lambda$ values are
equally distributed in $[0,0.05]$, while the other 20 $\lambda$ values
are equally distributed in $(0.05, 1]$. Therefore, in total 47
simulations are performed for all $\lambda$ values.
The free energy differences and the error estimates are calculated
by the Bennet's acceptance ratio method (BAR)~\cite{bennett1976efficient}. 
The simulation was
in system $\systemla$.

@@

The constant pressure molar heat capacity is defined in
a similar way, but at constant pressure
condition. 
It is calculated by estimating
the fluctuation of the enthalpy in an NPT simulation~\cite{wang2011existence}:
\begin{align}
  C_{p,m} = \frac{1}{k_BT^2 N} \langle ( H - \langle H\rangle)^2 \rangle,
\end{align}
where $ H$ is the enthalpy of the system,
defined by $H = \mathcal H + P\mathcal V$, and $\mathcal V$ is the instantaneous volume
of the system.
It has been shown that the statistical error in estimating
the fluctuations does not decrease with
respect to the system size~\cite{milchev1986fluctuations,ferrenberg1991statistical}.
Therefore, both constant volume and pressure molar heat capacities
are calculated from the simulation of the smallest system: $\systemsb$.

\subsection{Isothermal heat capacity}
The isothermal heat capacity is defined
by the normalized increment of volume due to an infiniesimal
decrement of pressure under constant temperature condition:
\begin{align}
  \kappa_T = - \frac 1V \Big(\frac{\partial V}{\partial P}\Big)_T
\end{align}
It can be calculated by estimating the fluctuation
of the instantaneous volume in an NPT simulation~\cite{wang2011existence}:
\begin{align}
  \kappa_T = \frac{1}{k_BT} \frac{\langle (\mathcal V - \langle \mathcal V\rangle)^2 \rangle}{\langle \mathcal V\rangle}.
\end{align}
Since it is also a fluctuation, the system $\systemsb$ is adopted for simulation.

\subsection{Thermal expansion coefficient}
The thermal expansion coefficient is defined by
the normalized increment of volume due to an infiniesimal
increment of temperature under the constant pressure condition:
\begin{align}
  \alpha_V = - \frac 1V \Big(\frac{\partial V}{\partial T}\Big)_P
\end{align}
It can be calculated by estimating the cross fluctuation
of the enthalpy and instantaneous volume  in an NPT simulation:
\begin{align}
  \alpha_V = \frac{1}{k_BT^2\langle \mathcal V\rangle} \langle (H - \langle H\rangle)\cdot(\mathcal V - \langle \mathcal V\rangle) \rangle
\end{align}
The system $\systemsb$ adopted for simulation.

@@

The dielectric constant can be estimated in a direct manner:
\begin{align}
  \eps = 1 + \frac{1}{3L^3 k_BT} ( \langle \vert \vect M\vert^2\rangle - \vert\langle \vect M\rangle\vert^2 )
\end{align}
where
\begin{align}
  \vect M = \sum_i\boldsymbol\mu_i = \sum_{\alpha\in i} q_\alpha\vect r_\alpha
\end{align}
is the total dipole moment of the system. Since it is a fluctuation, the simulated system is $\systemsb$.
% The results are listed in Tab.~\ref{tab:tmp2}. The dielectric constant calculated in system $\systemsb$ is reported in Table.~\ref{tab:thermo}.

@@

The Kirkwood G-factor is defined by~\cite{vanderSpoel2006origin}
\begin{align}
  G_k(r) =
  \Big\langle
  \frac 1N
  \sum_{i=1}^N \sum_{j, r_{ij} < r}
  \frac {\boldsymbol\mu_i \cdot \boldsymbol\mu_j}{\vert \boldsymbol\mu_i\vert \cdot \vert\boldsymbol\mu_j\vert}
  \Big\rangle,
\end{align}
where $r_{ij}$ denotes the oxygen-oxygen distance between two water
molecules.

\subsection{Diffusion constant}
The diffusion constant is calculated from the Einstein relation:
\begin{align}
  D = \lim_{t\rightarrow \infty}\frac {1}{6t} \langle \vert \vect r_i(t) - \vect r_i(0)\vert^2\rangle.
\end{align}
In practice, the mean-square-displacement $\langle\vert \vect r_i(t) - \vect
r_i(0)\vert^2\rangle$ is calculated, then the value is linearly fitted. The system for simulation is $\systemsb$.

@@

We investigate the convergence of the integral of the auto-correlation function w.r.t.~time:
\begin{align}
  I_\eta(T) = \frac{V}{k_BT}\int_0^T\langle P_{\alpha\beta}(0) P_{\alpha\beta}(t)\rangle\,dt,
\end{align}
and plot the function $I_\eta(T) $ for all method in
Fig.~\ref{fig:conv-vis}.  It is clear the integral for the RF method
converges at 6~ps, while the ZM method converges only in 2~ps, which
is consistent with the SPME result.  The value of ZM $l=1$ method is
different from the rest ZM method an SPME result.
The values listed  in Table~\ref{tab:thermo} are computed with $I_\eta(10\,\textrm{ps})$.
The
system for simulation is $\systemsb$.

>>>>

The equilibrium density of the system was determined by an NPT SPME
simulation.  In the NVT simulation the SPME electrostatic, computed
pressure ($-0.4\pm 1.5$~Bar) is reproducing the target pressure of
1.0~Bar.  The accuracy of ZM method, can be investigate by checking if
the NVT pressure is consistent with the target pressure.  In
Fig.~\ref{fig:pres-comp}, the pressure calculated by ZM method of
order $l=1,2,3$ and 4 are plotted against the cut-off radius.
All the  simulations are done with system $\systemla$.
It is concluded from the Figure that all ZM methods converges to the
target pressure at large enough cut-off radius.
In general, the higher order of ZM presents an larger error in the pressure calculation than
the lower order ZM method.
Higher order are also more sensitve in the  cut-off radius dependency.
ZM $l=1$ tends to under-estimate the pressure, while ZM $l=2,3$ and 4 tends to over-estimate.
The ZM method is also compared with
the RF method using the experimental dielectric constant $\varepsilon_{\textrm{rf}} = 80$, 
from which it is concluded that the accuracy of ZM in pressure calculation is much higher
than the RF method.
It should be noted that even with the largest system of 13824 water molecules,
the fluctuation (measured by standard deviation) of the pressure is roughly 100~Bar, which means that it is
very difficult achieve very accurate estimate for this quantity. In practice a pressure error of 20~Bar is
still acceptable.

\begin{figure}
  \centering
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l1.eps}
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l2.eps}\\
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l3.eps}
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l4.eps}
  \caption{The pressure convergence with respect to the cut-off radius
    for ZM methods with  splitting parameter from $0.0$ to $2.0\,\textrm{nm}^{-1}$.
    From left to right, up to down presensts the ZM method of order 1 to 4.
    The black solid line indicates the maganitude  of the correct pressure: 1~Bar.
    The error bars present the statistical uncertainty at 95\% confidence level.
  }
  \label{fig:pres-l1}
\end{figure}

The effect of using different splitting parameter on the convergence
of the pressure is investigate in Fig.~\ref{fig:pres-l1}.
In general, the higher order indicates less significant dependency on the splitting parameter.
In the case of $l=1$, the splitting parameter 0.0 is consistent with  $0.5\,\textrm{nm}^{-1}$, and
by using higher splitting parameter, the pressure shifts to the higher pressure side.
For $l=2,3$ and 4, the convergence is not sensitive to $\alpha$ upto  $1.0\,\textrm{nm}^{-1}$.
More importantly, for all orders,
the pressure does not converges to the desired value 1~Bar by using $\alpha = 1.5$ or $2.0\,\textrm{nm}^{-1}$.

@@

Other thermodynamic quantities including the chemical potential,
constant volume/pressure molar heat capacity, isothermal
compressibility and heat expansion coefficient are compared in
Tab.~\ref{tab:thermo}.
The ZM results are in good consistency with the SPME results, and out-perform
the RF method.
The splitting parameter does not have significant influence on these thermodynamic quantities.
\recheck{Discussion of the energy correction in the chemical potential calculation. TO be added later}

@@

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt/fig-rdf.eps}  
  \caption{The comparison of the center-of-mass radial distribution functions
    calculated by SPME, ZM and RF methods. The insert is a zoom-in of
    range 0.9 -- 1.5~nm. In the insert, the position of the cut-off
    (1.2~nm) is indicated by a vertical black line.
    For ZM method, the splitting parameter $\alpha$ was set to 0.0~$\textrm{nm}^{-1}$.
  }
  \label{fig:rdf}
\end{figure}

@@

\recheck{Shall we comment here on the fact that group cut-off produces smoother RDF? Add figures in Supplementary material}

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt/fig-rdf-l1-damp.eps}  
  \caption{
    The splitting parameter dependency in calculating the center-of-mass RDF for
    ZM method $l=1$. The insert is a zoom-in of
    range 0.9 -- 1.5~nm. In the insert, the position of the cut-off
    (1.2~nm) is indicated by a vertical black line.
  }
  \label{fig:rdf-damp}
\end{figure}

The artificial oscillation can be substantially improved by using larger splitting parameter $\alpha$.
In Fig.~\ref{fig:rdf-damp} presents the RDF of ZM $l=1$ with splitting parameter from 0.0 to 2.0~$\textrm{nm}^{-1}$.
A clear trend of better RDF by using larger splitting parameter is observed.
At $\alpha = 2.0\ \textrm{nm}^{-1}$ the RDF is almost overlapping with the SPME result.

\recheck{If we want $l=2,3$, add in supplementary material..}

@@

The dielectric constant computed by ZM method at different orders are listed in Tab.~\ref{tab:thermo},
and are consistent with the SPME computed results. In contrast, the RF method has poor accuracy in
the dielectric constant.

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt/fig-gkr.eps}
  \includegraphics[]{fig/result.nvt.small/fig-gkr-small.eps}
  \caption{The Kirkwood G-factor calculated for system $\systemlb$ (a) and $\systemmb$ (b).
    ZM method of orders $l=1,2,3$ and 4 with the cut-off 1.2~nm and splitting parameter $\alpha = 0.00\,\textrm{nm}^{-1}$ are plotted.
    The RF method is shown as an comparison, which uses a cut-off radius of 1.2~nm and dielectric constant of 80.
    The statistical uncertainty of the SPME method is presented at 95~\% confidence level with the red error bars.
  }
  \label{fig:gkr}
\end{figure}

As a measurement of the dipole correlation as a function of molecular distance, the Kirkwood-G factor
is an important qnantity of the water system.
It is well know that the G-factor is system size
dependent~\cite{vanderSpoel2006origin}, therefore we show the G-factor for
both  system $\systemlb$ in FIg.~\ref{fig:gkr}~(a) and a smaller system $\systemmb$
Fig.~\ref{fig:gkr} (b).  In the figures, the statistical uncertainty of
the SPME method is denoted by the red error bars. The statistical uncertainties of the
other methods are essentially of the same size as the SPME, so they are not presented in the figure for
clarity.  In the figure, ZM adaopts a cut-off radius of 1.2~nm and a splitting parameter $\alpha = 0.00\,\textrm{nm}^{-1}$.
RF uses a cut-off 1.2~nm and a dielectric constant of 80.

In both the system, the RF method is qualitatively
wrong. The accuracy of the ZM method is much higher than RF.
In the system of larger size ($\systemlb$), the ZM mehtod of $l=1$ is in consistency with the SPME mehtod. However
order $l=2,3$ and 4  present less but still acceptable accuracy.
In the system of smaller size ($\systemmb$), ZM method of all orders are consistent with the SPME result.
This means ZM performs better in calculated Kirkwood-G factor for smaller systems.
In both systems, ZM $l=1$ presents  an artificial
oscillation around the cut-off radius (1.2~nm), while this oscillation is
not obvious for $l\geq 2$.

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt/fig-gkr-conv.eps}
  \includegraphics[]{fig/result.nvt/fig-gkr-l3-damp.eps} 
  \caption{
    The cut-off (a) and spliting parameter (b) dependency in calculating the Kirkwood-G factor.
    The order of ZM method is $l=3$. In plot (a) the spliting parameter is $\alpha = 0.00\,\textrm{nm}^{-1}$.
    in plot (b) the cut-off radius is 1.2~nm.
    The statistical uncertainty is presented at 95~\% confidence level with the  error bars.
  }
  \label{fig:gkr-conv}
\end{figure}

The cut-off and spliting parameter dependency in calculating the Kirkwood-G factor
is presented in Fig.~\ref{fig:gkr-conv}  for ZM method of order $l=3$. With out surprising,
the Kirkwood-G factor converges to the SPME result when using increasing cut-off radius. 
However, unlike the case of calculating RDF, the accuracy Kirkwood-G factor is not improved
with increasing spliting parameter. This fact should be noticed in choosing parameters for
high accuracy calcualtion of Kirkwood-G factor.

\subsection{Dynamical quantities}

The dynamical quantities, diffusion constant and viscosity, are listed in Tab.~\ref{tab:thermo}.
The ZM method of order $l\geq 2$ can almost perfactly reproduce the SPME dynamical quantities.
The ZM order $l=1$ is of less accurate than higher orders, but is more accurate than the RF.
Using larger spliting parameters improves the accuracy, and the value $\alpha= 2.0\ \textrm{nm}^{-1}$ is
sugguested for calculating the dynamical quantities.

% \subsection{Damping effect}

% The convergence of the integrated auto-correlation function $I_\eta(T) $ is shown in Fig.~\ref{fig:damp-vis-l1}.
% \begin{figure}
%   \centering
%   \includegraphics[]{fig/result.tiny/fig-vis-l1-damp.eps}
%   \caption{The convergence of the integrated auto-correlation function $I_\eta(T) $. The error bars indicating 95\% confidence level are plotted with the SPME method. The different splitting parameters for ZM $l=1$ method are plotted}
%   \label{fig:damp-vis-l1}
% \end{figure}

% The RDF is plotted for different splitting parameters $\alpha$ for ZM $l=2$ and 3 in Fig.~\ref{fig:damp-rdf-l23}. Higher $\alpha$ reproduces the RDF better.
% The system was $L_1$.
% \begin{figure}[]
%   \centering
%   \includegraphics[width=0.49\textwidth]{fig/result.nvt/fig-rdf-l2-damp.eps}
%   \includegraphics[width=0.49\textwidth]{fig/result.nvt/fig-rdf-l3-damp.eps}
%   \caption{Radial distribution function for different splitting parameters $\alpha$ for ZM $l=2$ (left) and $l=3$ (right).}
%   \label{fig:damp-rdf-l23}
% \end{figure}

% The
% resulting slop is used to calculate the diffusion constant. The results are listed in Table~\ref{tab:thermo}.
% The ZM method $l\geq 2$ is consistent with the SPME
% result, while the RF method is off.

@@

\subsection{Radial distribution function}
The radial distribution function (RDF) $g(r)$ is a scalor function of distance $r$, which indicates the
probability of finding two atoms of distance $r$ apart.  It is a very
impotant equilibrium structure property, by which the X-ray scattering
intensity and a mount of thermodynamic properties can be computed.
In this work, we investigate the center-of-mass RDF in the water
system. The RDFs presented in the main text were computed in system $\systemlb$.

@@

\subsection{Pressure}
The pressure of the system calculated by the virial formulus:
\begin{align}
  P = \frac1{3V}
  \Bigg\langle
  \sum_{i=1}^{3N} \Big( m_i\vect v_i^2 + \vect r_i\cdot \vect F_i \Big)
  \Bigg\rangle,
\end{align}
where $V$ is the volume of the system. The presented pressures were computed in system $\systemla$.

@@

The chemical potential is defined by
\begin{align}
  \mu = \Big(\frac{\partial A}{\partial N}\Big)_{V,T},
\end{align}
where $A$ is the Helmholtz free energy.
In practice the excess chemical potential $\mu^\exc$, which is the
chemical potential abstracted by the kinetic contribution, is of
special interest. We calculate the excess chemical potential by the
thermodynamic integration (TI). In this approach, the interaction
between an inserted testing molecule and the system is
denotedd by $U_t(\lambda)$, where $\lambda$
is a coupling parameter in range $[0,1]$.
When $\lambda = 1$ the testing molecule is fully coupled to the system
with the TIP3P interaction. When $\lambda=0$, the testing molecule
is decoupled with the system.
Therefore the
Hamiltonian of the system is function of the coupling parameter. The
excess chemical is calculated by the free energy difference between
the fully coupled and decoupled states:
\begin{align}
  \mu^\exc = \int_0^1 \Big\langle \frac{\partial U_t(\lambda)}{\partial \lambda} \Big\rangle_\lambda d\lambda
\end{align}
This integral is usually calculated by numerical integration, which
equally paritions the range of $\lambda$, simulates the system at each
discretized $\lambda$, and computes chemical potential by trapezoidal formulus.
Here for the sake of accuracy, 
we adopt a two step coupling approach: firstly couple the van der Waals
interaction of the testing particle with the rest of the system (stage vdw),
and then couple the electrostatic interaction (stage ele).
During stage vdw, 21 $\lambda$ values are equally
distributed in $[0,1]$.  During stage ele, 6 $\lambda$ values are
equally distributed in $[0,0.05]$, while the other 20 $\lambda$ values
are equally distributed in $(0.05, 1]$. Therefore, in total 47
simulations are performed for all $\lambda$ values.
The free energy differences and the error estimates are calculated
by the Bennet's acceptance ratio method (BAR)~\cite{bennett1976efficient}. 
The simulations were performed with system $\systemla$.

@@

The constant pressure molar heat capacity is defined in
a similar way, but at constant pressure
condition. 
It is calculated by estimating
the fluctuation of the enthalpy in an NPT simulation~\cite{wang2011existence}:
\begin{align}
  C_{p,m} = \frac{1}{k_BT^2 N} \langle ( H - \langle H\rangle)^2 \rangle,
\end{align}
where $ H$ is the enthalpy of the system,
defined by $H = \mathcal H + P\mathcal V$, and $\mathcal V$ is the instantaneous volume
of the system.
It has been shown that the statistical error in estimating
the fluctuations does not decrease with
respect to the system size~\cite{milchev1986fluctuations,ferrenberg1991statistical}.
Therefore, both constant volume and pressure molar heat capacities
were calculated from the simulation of the smallest system: $\systemsb$.

\subsection{Isothermal heat capacity}
The isothermal heat capacity is defined
by the normalized increment of volume due to an infiniesimal
decrement of pressure under constant temperature condition:
\begin{align}
  \kappa_T = - \frac 1V \Big(\frac{\partial V}{\partial P}\Big)_T
\end{align}
It can be calculated by estimating the fluctuation
of the instantaneous volume in an NPT simulation~\cite{wang2011existence}:
\begin{align}
  \kappa_T = \frac{1}{k_BT} \frac{\langle (\mathcal V - \langle \mathcal V\rangle)^2 \rangle}{\langle \mathcal V\rangle}.
\end{align}
Since it is also a fluctuation, the system $\systemsb$ was adopted for simulation.

\subsection{Thermal expansion coefficient}
The thermal expansion coefficient is defined by
the normalized increment of volume due to an infiniesimal
increment of temperature under the constant pressure condition:
\begin{align}
  \alpha_V = - \frac 1V \Big(\frac{\partial V}{\partial T}\Big)_P
\end{align}
It can be calculated by estimating the cross fluctuation
of the enthalpy and instantaneous volume  in an NPT simulation:
\begin{align}
  \alpha_V = \frac{1}{k_BT^2\langle \mathcal V\rangle} \langle (H - \langle H\rangle)\cdot(\mathcal V - \langle \mathcal V\rangle) \rangle
\end{align}
The system $\systemsb$ was adopted for simulation.

@@

The dielectric constant can be estimated in a direct manner:
\begin{align}
  \eps = 1 + \frac{1}{3L^3 k_BT} ( \langle \vert \vect M\vert^2\rangle - \vert\langle \vect M\rangle\vert^2 )
\end{align}
where
\begin{align}
  \vect M = \sum_i\boldsymbol\mu_i = \sum_{\alpha\in i} q_\alpha\vect r_\alpha
\end{align}
is the total dipole moment of the system. Since it is a fluctuation, the simulated system was $\systemsb$.
% The results are listed in Tab.~\ref{tab:tmp2}. The dielectric constant calculated in system $\systemsb$ is reported in Table.~\ref{tab:thermo}.

@@

The Kirkwood G-factor is defined by~\cite{vanderSpoel2006origin}
\begin{align}
  G_k(r) =
  \Big\langle
  \frac 1N
  \sum_{i=1}^N \sum_{j, r_{ij} < r}
  \frac {\boldsymbol\mu_i \cdot \boldsymbol\mu_j}{\vert \boldsymbol\mu_i\vert \cdot \vert\boldsymbol\mu_j\vert}
  \Big\rangle,
\end{align}
where $r_{ij}$ denotes the oxygen-oxygen distance between two water
molecules.  The system for computing this quantity is $\systemlb$.

\subsection{Diffusion constant}
The diffusion constant is calculated from the Einstein relation:
\begin{align}
  D = \lim_{t\rightarrow \infty}\frac {1}{6t} \langle \vert \vect r_i(t) - \vect r_i(0)\vert^2\rangle.
\end{align}
In practice, the mean-square-displacement $\langle\vert \vect r_i(t) - \vect
r_i(0)\vert^2\rangle$ is calculated, then the value is linearly fitted. The system for simulation was $\systemsb$.

@@

We investigate the convergence of the integral of the auto-correlation function w.r.t.~time:
\begin{align}
  I_\eta(T) = \frac{V}{k_BT}\int_0^T\langle P_{\alpha\beta}(0) P_{\alpha\beta}(t)\rangle\,dt,
\end{align}
and plot the function $I_\eta(T) $ for all method in
Fig.~\ref{fig:conv-vis}.  It is clear the integral for the RF method
converges at 6~ps, while the ZM method converges only in 2~ps, which
is consistent with the SPME result.  The value of ZM $l=1$ method is
different from the rest ZM method an SPME result.
The values listed  in Table~\ref{tab:thermo} are computed with $I_\eta(10\,\textrm{ps})$.
The
system for simulation was $\systemsb$.
